% --- ABSTRACT ---
\begin{center}
    \textbf{Abstract}
\end{center}

\noindent This Final Degree Project details the design, development, and implementation of a high-performance rendering engine using the Vulkan API, with a specialized focus on the simulation and visualization of complex particle systems in real time. The core of the project is based on the creation of a parallel computing architecture that allows for the simulation of atmospheric and dynamic phenomena such as fire, smoke, and water through the intensive use of Compute Shaders. This technical approach transfers the entire computational load of the particle lifecycle —including position, color, opacity, and physical transformation updates— entirely to the GPU, effectively eliminating the communication bottlenecks between the CPU and video memory typically found in higher-level APIs. At the implementation level, the engine stands out for its explicit management of hardware resources through a modular structure that integrates a procedural computation pipeline, advanced synchronization systems via memory barriers (VkBufferMemoryBarrier), and volumetric rendering techniques based on Point Sprites and Alpha Blending. Furthermore, the application integrates a complete three-dimensional graphical environment with ADS (Ambient, Diffuse, Specular) lighting, optimized 3D model loading, and a dynamic Skybox system. The modular nature of the engine facilitates simple extensibility, while optimized memory management through descriptors and local buffers ensures low data access latency. The results obtained validate Vulkan's high efficiency in processing over one million concurrent particles while maintaining a stable frame rate above 60 FPS on mid-range hardware, allowing for high aesthetic parameterization. This positions the developed tool as a solid, scalable, and robust foundation for interactive simulations and advanced applications within the video game industry.

\vspace{0.5cm}
\textit{Keywords:} vulkan, compute shaders, particle systems, real-time, gpu

\vspace{0.5cm}
\textit{Specific competencies:} 
In this work, the following specific competencies have been developed:
\begin{itemize}
    \item Ability to develop and evaluate interactive systems and complex information presentation and their application to solving human-computer interaction design problems.
\end{itemize}
